Quasicharacters and the function-sheaf dictionary

### 2014-09-04 crelle-dictionary.tex
Finished working through paper

\begin{proof}
We can describe $X^{\Weil{}}$ as the kernel of the map $X \xrightarrow{\Frob{}-1} X$;
let $Y = (\Frob{}-1)X$ be the augmentation ideal.  Dualizing the sequence
\[
 0 \to X^{\Weil{}} \to X \to Y \to 0
\]
yields
\[
 0 \to Y^* \to X^* \to (X^{\Weil{}})^* \to \Ext^1_\ZZ(Y, \EEx).
\]
Since $\Ext^1_\ZZ(-,\EEx)$ vanishes, we get a natural isomorphism from the cokernel of $Y^* \xrightarrow{\Frob{}-1} X^*$ to $(X^{\Weil{}})^*$.
\todo{Can we use this sort or argument to prove $(X_{\Weil{}})^* \iso (X^*)^{\Weil{}}$? If so, let's re-state the $\Aut(\qcs{L})$ result as $\Hh^1(\pi_0(\bG),\EEx)^{\Weil{}}$ since the iso classes are $\Hh^2(\pi_0(\bG),\EEx)^{\Weil{}}$.}
\end{proof}

@@

\begin{proposition}\label{prop:connected}
 If $G$ is a connected, commutative algebraic group over $\Fq$ then:
 \begin{enumerate}
 \item $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$ is an isomorphism of groups;
 \item every quasicharacter sheaf on $G$ is isomorphic to one defined by a discrete isogeny;%as in Proposition~\ref{prop:finite};
 \item $\Aut(\qcs{L}) = 1$, for all quasicharacter sheaves $\qcs{L}$ on $G$.
%  \item for all quasicharacter sheaves $\qcs{L}$, $\qcs{L}'$ on $G$ and for every $\rho \in \Hom(\qcs{L},\qcs{L}')$, either $\rho$ is trivial or an isomorphism.
 \end{enumerate}
 \end{proposition}
\begin{proof}
In Lemma~\ref{lem:bounded_connected} we saw that every quasicharacter sheaf $\qcs{L}$ on $\bG$ descends to $G$; let $\qcs{E}$ be an object of $\bQC(G)$ for which $B_G(\qcs{E}) \iso \qcs{L}$.
 Since the functor $B_G : \bQC(G) \to \QC(G)$ is full and faithful, $\Aut(\qcs{L}) = \Aut(\qcs{E})$.
From here, Deligne's function-sheaf dictionary for connected commutative algebraic groups over finite fields, as in \cite{deligne:SGA4.5}*{Sommes trig.} or \cite{laumon:87a}*{1.1.3}, gives us all we need, as we briefly recall. %\todo{Fix citelist} %\citelist{\cite{deligne:SGA4.5}*{Sommes trig.},\cite{laumon:87a}*{1.1.3}}

As in the proof of Proposition~\ref{prop:finite}, use \'etale descent to see that pullback by the Lang isogeny $\Lang : G\to G$ defines an equivalence of categories from local systems on $G$ and $G(\Fq)$-equivariant local systems on $G$. 
Under this equivalence, local systems $\qcs{E}$ on $G$ arising from objects in $\bQC(G)$ are matched with $G(\Fq)$-equivariant constant local systems of rank-one on $G$, and therefore with one-dimensional representations of $G(\Fq)$. 
In the same way, pullback along the isogeny $\Lang\times\Lang : G\times G\to G\times G$ matches the extra structure $\mu_\qcs{E} : m^*\qcs{E} \to \qcs{E}\boxtimes\qcs{E}$ with an isomorphism $m^*V \to V\boxtimes V$ of one-dimensional representations of $G(\Fq)\times G(\Fq)$, which is exactly an isomorphism $V \to V\otimes V$ of one-dimensional representations, which is exactly the choice of an isomorphism $V\iso \EE$.
In this way we see that $\bQC(G)$ is equivalent to the category of characters of $G(\Fq)$.
A simple calculation on stalks at geometric points ${\bar g}$ above $\Fq$-rational points $g \in G(\Fq)$ reveals that, if $\qcs{E}$ matches $\psi : G(\Fq)\to \EEx$ under this equivalence, then the action of Frobenius on $\qcs{E}{\bar g}$ is multiplication by $\psi(g)^{-1}$.
In other words, for every $\qcs{E}$ in $\bQC(G)$, the trace of $\Lang^*\qcs{E}$, as a representation of $G(\Fq)$, is $\trFrob{\qcs{E}}^{-1}$.
This proves parts (i) and (ii) of Proposition~\ref{prop:connected}.

To finish the proof of part (iii), suppose $\Lang^*\qcs{E} = V$ with isomorphism $V \to V\otimes V$ and observe that the equivalence above establishes a bijection between $\Aut(\qcs{E})$ and the group of automorphisms of $\rho : V\to V$ for which 
\[
\begin{tikzcd}
\arrow{d}{} V \arrow{r}{\rho} & V\arrow{d}{}\\
V\otimes V \arrow{r}{\rho\otimes \rho} & V\otimes V
\end{tikzcd}
\]
commutes. 
Since the only such isomorphism $\rho$ is $\id_V$, it follows that $\Aut(\qcs{E}) = 1$, completing the proof.
\end{proof}

%\begin{remark}
We have just seen that, for a connected commutative algebraic group $G$ over $\Fq$, the category of quasicharacter sheaves on $G$ is equivalent to the category of one-dimensional representations $V$ of $G(\Fq)$ equipped with an isomorphism $V\iso \EE$, and therefore equivalent to the category of characters $\psi$ of $G(\Fq)$.
We have also just seen that if the character of $\Lang^*\qcs{E}$ is $\psi$ then the canonical isomorphism $m^*\psi \iso \psi \boxtimes \psi$ determines the isomorphism $\mu_\qcs{E} : \qcs{E} \to \qcs{E}\boxtimes\qcs{E}$ which is part of the data that determines a quasicharacter sheaf on $G$.
That leads to a perspective on the function-sheaf dictionary common in the literature in which one considers one-dimensional local systems $\qcs{E}$ on $G$ for which \emph{there exists} an isomorphism $m^*\qcs{E} \iso \qcs{E} \boxtimes\qcs{E}$; see \cite{kamgarpour:09a}*{Intro}, for example.
As a slight variation, one may also consider one-dimensional local systems $\gqcs{L}$ on $\bG$ for which \emph{there exists} an isomorphism $\Frob{G}^*\gqcs{L} \iso \gqcs{L}$ and an isomorphism $\bm^*\gqcs{L} \iso \gqcs{L} \boxtimes\gqcs{L}$.
%One may show that all such $\gqcs{L}$ are determined by extensions of $\bG$ by $\EEx$ which are fixed by Frobenius and that the trace of Frobenius provides a canonical isomorphism from $\Hh^2(\bG,\EEx)^{\Frob{}}$ to $G(\Fq)^*$; {\it cf.\ }
%One may show that all such $\gqcs{L}$ are determined by extensions of $\bG$ by $\EEx$ which are fixed by Frobenius and that the trace of Frobenius provides a canonical isomorphism from $\Hh^2(\bG,\EEx)^{\Frob{}}$ to $G(\Fq)^*$; {\it cf.\ } \cite{kamgarpour:09a}*{Prop 2.15}.

Although the category of quasicharacter sheaves on $G$ does specialize to that common notion when $G$ is connected and of finite type, the common notion is not appropriate if one wishes to extend the dictionary to smooth commutative group schemes. 
In particular, as we have already seen in Section~\ref{sec:etale}, for a given quasicharacter sheaf $\qcs{L} = (\gqcs{L},\mu,\phi)$ there may be a different isomorphism $\mu' : \bm^*\gqcs{L} \to \gqcs{L}\boxtimes\gqcs{L}$ for which $\qcs{L}'=(\gqcs{L},\mu',\phi)$ is also a quasicharacter sheaf on $G$, not isomorphic to $\qcs{L}$ in $\QC(G)$.
Indeed, Proposition~\ref{prop:etale-iso} shows that when $G$ is \'etale, $\Hh^2(\bG,\EEx)^{\Weil{}}$ measures the possibilities for $\mu'$.
On \'etale group schemes, the common perspective on the function-sheaf dictionary leads only to the part of category of quasicharacter sheaves that is invisible to the trace of Frobenius.

We will see in Section~\ref{sec:main} that, for any smooth commutative group scheme $G$, the part of the category of quasicharacter sheaves that is invisible to the trace of Frobenius is measured by $\Hh^2(\pi_0(\bG),\EEx)^{\Weil{}}$.
%\end{remark}

@@

Consider the short exact sequence
defining the component group scheme for $G$:
\begin{equation}\label{eq:pi0}
\begin{tikzcd}
0 \rar & G^0 \arrow{r}{\iota_0} & G \arrow{r}{\pi_0} & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is a \'etale commutative group scheme -- and thus smooth --
Lemma~\ref{lem:pullback} implies that \eqref{eq:pi0} defines the sequence of functors
\begin{equation}\label{eq:pi1}
\begin{tikzcd}
\QC(0) \rar & \QC(\pi_0(G)) \arrow{r}{\pi_0^*} & \QC(G) \arrow{r}{\iota_0^*} & \QC(G^0) \rar & \QC(0)
\end{tikzcd}
\end{equation}
and therefore, after passing to isomorphism classes, the sequence of abelian groups
\begin{equation}\label{eq:pi2}
\begin{tikzcd}
0 \rar &
\QCiso{\pi_0(G)} \arrow{r}{\pi_0^*} & \QCiso{G} \arrow{r}{\iota_0^*} & \QCiso{G^0} \rar & 0.
\end{tikzcd}
\end{equation}
 Note that we found the groups $\QCiso{\pi_0(G)}$ and $\QCiso{G^0}$
in Sections~\ref{ssec:SandT} and \ref{ssec:connected}, respectively.
We will shortly see that \eqref{eq:pi2} is exact.

@@

\begin{proof}
Let $\pi: B \to G^0$ be a discrete isogeny; set $A \ceq \ker \pi$.
  We will find a discrete isogeny $f: H\to G$
  such that that $H^0 = B$, $f^0 =\pi$ and
  $\pi_0(f) : \pi_0(H)\to \pi_0(G)$ is an isomorphism of component
  groups.  Namely, we will fit $\pi$ into the following diagram,
  \begin{equation}\label{extension-diagram}
  \begin{tikzcd}
  A \arrow[equal]{r} \dar & A \dar \\
  B \rar \dar[swap]{\pi} & H \rar \dar[swap]{f} & \pi_0(H) \arrow{d}[below,rotate=90]{\sim}[swap]{\pi_0(f)} \\
  G^0 \rar & G \rar & \pi_0(G),
  \end{tikzcd}
  \end{equation}
  where all rows and columns are exact and all maps are defined over
  $\Fq$.  We will do so by passing back and forth between group
  schemes over $\Fq$ and their $\bFq$-points.

@@

We now wish to use the apply the results of Section~\ref{ssec:connected} to the identity component of $G$; to do so, we must know that the identity component of $G$ is actually an algebraic group over $\Fq$.

\begin{lemma} \label{lem:G0alg-grp}
If $G$ is a commutative smooth group scheme over $\Fq$ then its identity component, $G^0$, is a connected algebraic group over $\Fq$.
\end{lemma}
\begin{proof}
 Since $G$ is a smooth group scheme over $\Fq$, its
 identity component $G^0$ of $G$ is a connected smooth,
 group scheme of finite type over $\Fq$, reduced over some finite extension of $\Fq$
 \cite{vdGeer-Moonen:AbelianVarieties}*{3.17}.
 Since $\Fq$ is a finite field and hence perfect, $G^0$ is actually reduced over $\Fq$
 \cite{EGAIV2}*{Prop 6.4.1}.  Since every group scheme over a field is separated
 \cite{vdGeer-Moonen:AbelianVarieties}*{3.12},
 it follows that $G^0$ is a connected algebraic group.
\end{proof}

@@

\begin{lemma}\label{lem:extension}
The group homomorphism $\pi_0^*: \QCiso{\pi_0(G)} \to \QCiso{G}$ is injective.
\end{lemma}
\begin{proof}
Let $\qcs{L}$ be a quasicharacter sheaf on $\pi_0(G)$ and let $\rho : \pi_0^*\qcs{L} \to (\EE)_{G}$ be an isomorphism in $\QC(G)$. 
For each $x\in \pi_0(\bG)$, set $\bG^x \ceq \pi_0^{-1}(x)$.
The restriction $\pi_0^*\gqcs{L}\vert_{\bG^x}$ is the constant sheaf $(\gqcs{L}_x)_{\bG^x}$ so the isomorphism $\brho_{\bG^x} : (\gqcs{L}_x)_{\bG^x} \to (\EE)_{\bG^x}$ determines an isomorphism $\brho_x : \gqcs{L}_x \to (\EE)_x$. The collection $\{ \brho_x \tq x\in \pi_0(\bG) \}$ determines an isomorphism $\qcs{L} \to (\EE)_{\pi_0(G)}$  in $\QC(\pi_0(G))$.
\end{proof}

@@

\begin{remark}
Although $\TrFrob{\pi_0(G)}$ is split and $\TrFrob{G^0}$ is an isomorphism, as examples show, it does not follow that $\TrFrob{G}$ is split. 
\end{remark}

@@

In the opposite direction, let $\Nm : G(k') \to G(k)$ be the norm map and consider the group homomorphism:
\begin{align*}
\Nm^* : G(k)^* &\to G(k')^* \\
\chi &\mapsto \chi\circ \Nm.
\end{align*}
We can also interpret this operation on characters in terms of quasicharacter sheaves.

If $\qcs{L} \ceq (\gqcs{L}, \mu, \phi)$ is a quasicharacter sheaf on $G$, we define
%a quasicharacter sheaf 
$\qcs{L}' \ceq (\gqcs{L}, \mu, \phi_{k'})$ on the base change
$G_{k'}$ of $G$ to $k'$ by setting
\[
\phi_{k'} \ceq \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi).
\]
The commutativity of the diagram (QC.3) for $\phi_{k'}$
follows from the fact that $\Frob{G_{k'}} = \Frob{G}^n$.
Note that we may also think about the construction of $\phi_{k'}$ from $\phi$
as taking the action $\varphi$ of the Weil group for $k$%, $\Weil{k}$, 
on $\gqcs{L}$
defined in Section~\ref{ssec:category} and restricting it to the Weil group for $k'$.
%, $\Weil{k'}$.

@@

\begin{proof}
Let $\qcs{L} \ceq (\gqcs{L}, \mu, \phi) \in \QC(G)$ and write $F$ for $\Frob{G}$.  For any $x \in G(k')$,
we may compute the value of $t_{G_{k'}}(\nu_{k'/k}\qcs{L})(x)= t_{\nu_{k'/k}\qcs{L}}(x)$ as the trace of $\phi_{k'}$ on $\gqcs{L}_x$,
and the value of $\Nm^*(\TrFrob{G}(\qcs{L}))(x)$ as the trace of $\phi$ on $\gqcs{L}_{\Nm(x)}$.
Applying \ref{QC.3} to the stalk of $\gqcs{L}^{\boxtimes n}$ at the point $(x, \Frob{}(x), \ldots, \Frob{}^{n-1}(x))$ yields a diagram
\[
\begin{tikzcd}
\gqcs{L}_{\Nm(x)} \rar \dar{\phi_{\Nm(x)}} & \gqcs{L}_{F(x)} \otimes \gqcs{L}_{F^2(x)} \otimes \cdots \otimes \gqcs{L}_x
\dar{\phi_x \otimes (F^*\phi)_x \otimes \cdots \otimes ((F^{n-1})^*\phi)_x} \\
\gqcs{L}_{\Nm(x)} \rar & \gqcs{L}_x \otimes \gqcs{L}_{F(x)} \otimes \cdots \otimes \gqcs{L}_{F^{n-1}(x)}.
\end{tikzcd}
\]
Choose a basis vector $v$ for $\gqcs{L}_{\Nm(x)}$ and write $v_0 \otimes v_1 \otimes \cdots \otimes v_{n-1}$ for the image of $v$ under the
bottom map,
for $v_i \in \gqcs{L}_{\Frob{}^i(x)}$.  By \ref{QC.2}, $v$ maps to
$v_1 \otimes v_2 \otimes \cdots \otimes v_0$ along the top of the diagram.
Let $\alpha_i \in \EEx$ represent $((F^i)^*\phi)_x$ with respect to these bases and let $\alpha$ be
the trace of $\phi_{\Nm(x)}$.  We may now equate the trace $\alpha$ of $\phi$ on $\gqcs{L}_{\Nm(x)}$
with the product $\alpha_0 \cdots \alpha_{n-1}$, which is the trace of $\phi_{k'}$ on $\gqcs{L}_x$.
\end{proof}

Finally, let $G'$ be a smooth commutative group scheme over $k'$.
We explain how to geometrize the canonical isomorphism between characters of $G'(k')$ and of $(\Res_{k'/k}G')(k)$.
We may decompose the base change $(\Res_{k'/k}G')_{k'}$ of $\Res_{k'/k}G'$ to $k'$
into a product of copies of $G'$, indexed by elements of $\Gal(k'/k)$:
\[
(\Res_{k'/k}G')_{k'} \cong \prod_{\Gal(k'/k)} G'.
\]
Since products and coproducts agree for group schemes we have a natural inclusion of $k'$-schemes
\[
G' \hookrightarrow (\Res_{k'/k}G')_{k'}
\]
mapping $G'$ into the summand corresponding to $1 \in \Gal(k'/k)$.  Composing $\nu_{k'/k}$
from Proposition~\ref{prop:basechange} with pullback along this map yields a functor
\[
\rho : \QC(\Res_{k'/k}G') \to \QC(G').
\]

>>>>

\begin{proof}
We can describe $X^{\Weil{}}$ as the kernel of the map $X \xrightarrow{\Frob{}-1} X$;
let $Y = (\Frob{}-1)X$ be the augmentation ideal.  Dualizing the sequence
\[
 0 \to X^{\Weil{}} \to X \to Y \to 0
\]
yields
\[
 0 \to Y^* \to X^* \to (X^{\Weil{}})^* \to \Ext^1_\ZZ(Y, \EEx).
\]
Since $\Ext^1_\ZZ(-,\EEx)$ vanishes, we get a natural isomorphism from the cokernel of $Y^* \xrightarrow{\Frob{}-1} X^*$ to $(X^{\Weil{}})^*$.
\end{proof}

@@

Using this equivalence of categories, we may give a good description of $\QC(G)$ when $G$ is connected and finite type.

\begin{proposition}\label{prop:connected}
 If $G$ is a connected, commutative algebraic group over $\Fq$ then:
 \begin{enumerate}
 \item $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$ is an isomorphism of groups;
 \item every quasicharacter sheaf on $G$ is isomorphic to one defined by a discrete isogeny;%as in Proposition~\ref{prop:finite};
 \item $\Aut(\qcs{L}) = 1$, for all quasicharacter sheaves $\qcs{L}$ on $G$.
%  \item for all quasicharacter sheaves $\qcs{L}$, $\qcs{L}'$ on $G$ and for every $\rho \in \Hom(\qcs{L},\qcs{L}')$, either $\rho$ is trivial or an isomorphism.
 \end{enumerate}
 \end{proposition}
\begin{proof}
In Lemma~\ref{lem:bounded_connected}, we saw that every quasicharacter sheaf $\qcs{L}$ on $\bG$ descends to $G$; let $\qcs{E}$ be an object of $\bQC(G)$ for which $B_G(\qcs{E}) \iso \qcs{L}$.
 Since the functor $B_G : \bQC(G) \to \QC(G)$ is full and faithful, $\Aut(\qcs{L}) = \Aut(\qcs{E})$.
From here, Deligne's function-sheaf dictionary for connected commutative algebraic groups over finite fields, as in \cite{deligne:SGA4.5}*{Sommes trig.} or \cite{laumon:87a}*{1.1.3}, gives us all we need, as we briefly recall. %\todo{Fix citelist} %\citelist{\cite{deligne:SGA4.5}*{Sommes trig.},\cite{laumon:87a}*{1.1.3}}

As in the proof of Proposition~\ref{prop:finite}, use \'etale descent to see that pullback by the Lang isogeny $\Lang : G\to G$ defines an equivalence of categories between local systems on $G$ and $G(\Fq)$-equivariant local systems on $G$. 
Under this equivalence, local systems $\qcs{E}$ on $G$ arising from objects in $\bQC(G)$ are matched with $G(\Fq)$-equivariant constant local systems of rank-one on $G$, and therefore with one-dimensional representations of $G(\Fq)$. 
In the same way, pullback along the isogeny $\Lang\times\Lang : G\times G\to G\times G$ matches the extra structure $\mu_\qcs{E} : m^*\qcs{E} \to \qcs{E}\boxtimes\qcs{E}$ with an isomorphism $m^*V \to V\boxtimes V$ of one-dimensional representations of $G(\Fq)\times G(\Fq)$, which is exactly an isomorphism $V \to V\otimes V$ of one-dimensional representations, which is exactly the choice of an isomorphism $V\iso \EE$.
We see that $\bQC(G)$ is equivalent to the category of characters of $G(\Fq)$.
Let $\bg$ be a geometric point above $g \in G(\Fq)$.  If $\qcs{E}$ matches $\psi : G(\Fq)\to \EEx$ under this equivalence, a simple calculation on stalks reveals that the action of Frobenius on $\qcs{E}_{\bar g}$ is multiplication by $\psi(g)^{-1}$.
In other words, for every $\qcs{E}$ in $\bQC(G)$, the trace of $\Lang^*\qcs{E}$ is $\trFrob{\qcs{E}}^{-1}$ as a representation of $G(\Fq)$, proving parts (i) and (ii).

To finish the proof of part (iii), suppose $\Lang^*\qcs{E} = V$ with isomorphism $V \to V\otimes V$.  Observe that the equivalence above establishes a bijection between $\Aut(\qcs{E})$ and the group of automorphisms of $\rho : V\to V$ for which 
\[
\begin{tikzcd}
\arrow{d}{} V \arrow{r}{\rho} & V\arrow{d}{}\\
V\otimes V \arrow{r}{\rho\otimes \rho} & V\otimes V
\end{tikzcd}
\]
commutes. 
Since the only such isomorphism $\rho$ is $\id_V$, it follows that $\Aut(\qcs{E}) = 1$, completing the proof.
\end{proof}

%\begin{remark}
We have just seen that, for a connected commutative algebraic group $G$ over $\Fq$, the category of quasicharacter sheaves on $G$ is equivalent to the category of one-dimensional representations $V$ of $G(\Fq)$ equipped with an isomorphism $V\iso \EE$, and therefore equivalent to the category of characters $\psi$ of $G(\Fq)$.
We have also just seen that if the character of $\Lang^*\qcs{E}$ is $\psi$ then the canonical isomorphism $m^*\psi \iso \psi \boxtimes \psi$ determines the isomorphism $\mu_\qcs{E} : \qcs{E} \to \qcs{E}\boxtimes\qcs{E}$.
This fact leads to a perspective on the function-sheaf dictionary common in the literature in which one considers one-dimensional local systems $\qcs{E}$ on $G$ for which \emph{there exists} an isomorphism $m^*\qcs{E} \iso \qcs{E} \boxtimes\qcs{E}$ \cite{kamgarpour:09a}*{Intro}.
As a slight variation, one may also consider one-dimensional local systems $\gqcs{L}$ on $\bG$ for which \emph{there exists} an isomorphism $\Frob{G}^*\gqcs{L} \iso \gqcs{L}$ and an isomorphism $\bm^*\gqcs{L} \iso \gqcs{L} \boxtimes\gqcs{L}$.
%One may show that all such $\gqcs{L}$ are determined by extensions of $\bG$ by $\EEx$ which are fixed by Frobenius and that the trace of Frobenius provides a canonical isomorphism from $\Hh^2(\bG,\EEx)^{\Frob{}}$ to $G(\Fq)^*$; {\it cf.\ }
%One may show that all such $\gqcs{L}$ are determined by extensions of $\bG$ by $\EEx$ which are fixed by Frobenius and that the trace of Frobenius provides a canonical isomorphism from $\Hh^2(\bG,\EEx)^{\Frob{}}$ to $G(\Fq)^*$; {\it cf.\ } \cite{kamgarpour:09a}*{Prop 2.15}.

Although the category of quasicharacter sheaves on $G$ specializes to the one described above when $G$ is connected and of finite type, this description is not sufficient when extending the dictionary to smooth commutative group schemes. 
In particular, for a given $\gqcs{L}$ and $\phi$ there may be many $\mu$ so that $(\gqcs{L},\mu,\phi)$ is a quasicharacter sheaf.  For \'etale $G$, Proposition~\ref{prop:etale-iso} shows that $\Hh^2(\bG,\EEx)^{\Weil{}}$ measures the possibilities for $\mu$.  We will see in Section~\ref{sec:main} that $\Hh^2(\pi_0(\bG),\EEx)^{\Weil{}}$ plays an analogous role for general smooth commutative group schemes $G$.

@@

Consider the short exact sequence
defining the component group scheme for $G$:
\begin{equation}\label{eq:pi0}
\begin{tikzcd}
0 \rar & G^0 \arrow{r}{\iota_0} & G \arrow{r}{\pi_0} & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is an \'etale commutative group scheme -- and thus smooth --
Lemma~\ref{lem:pullback} implies that \eqref{eq:pi0} defines a sequence of functors
\begin{equation}\label{eq:pi1}
\begin{tikzcd}
\QC(0) \rar & \QC(\pi_0(G)) \arrow{r}{\pi_0^*} & \QC(G) \arrow{r}{\iota_0^*} & \QC(G^0) \rar & \QC(0)
\end{tikzcd}
\end{equation}
and therefore, after passing to isomorphism classes, a sequence of abelian groups
\begin{equation}\label{eq:pi2}
\begin{tikzcd}
0 \rar &
\QCiso{\pi_0(G)} \arrow{r}{\pi_0^*} & \QCiso{G} \arrow{r}{\iota_0^*} & \QCiso{G^0} \rar & 0.
\end{tikzcd}
\end{equation}
 Note that we found the groups $\QCiso{\pi_0(G)}$ and $\QCiso{G^0}$
in Sections~\ref{ssec:SandT} and \ref{ssec:connected}, respectively.
We will shortly see that \eqref{eq:pi2} is exact.

@@

\begin{proof}
Let $\pi: B \to G^0$ be a discrete isogeny, and set $A \ceq \ker \pi$.
  We will find a discrete isogeny $f: H\to G$
  such that that $H^0 = B$, $f^0 =\pi$ and
  $\pi_0(f) : \pi_0(H)\to \pi_0(G)$ is an isomorphism of component
  groups.  Namely, we will fit $\pi$ into the following diagram,
  \begin{equation}\label{extension-diagram}
  \begin{tikzcd}
  A \arrow[equal]{r} \dar & A \dar \\
  B \rar \dar[swap]{\pi} & H \rar \dar[swap]{f} & \pi_0(H) \arrow{d}[below,rotate=90]{\sim}[swap]{\pi_0(f)} \\
  G^0 \rar & G \rar & \pi_0(G),
  \end{tikzcd}
  \end{equation}
  where all rows and columns are exact and all maps are defined over
  $\Fq$.  We will do so by passing back and forth between group
  schemes over $\Fq$ and their $\bFq$-points.

@@

We now wish to apply the results of Section~\ref{ssec:connected} to the identity component of $G$; to do so, we must know that the identity component of $G$ is actually an algebraic group over $\Fq$.

\begin{lemma} \label{lem:G0alg-grp}
If $G$ is a commutative smooth group scheme over $\Fq$ then its identity component, $G^0$, is a connected algebraic group over $\Fq$.
\end{lemma}
\begin{proof}
 Since $G$ is a smooth group scheme over $\Fq$, its
 identity component $G^0$ is a connected smooth,
 group scheme of finite type over $\Fq$, reduced over some finite extension of $\Fq$
 \cite{vdGeer-Moonen:AbelianVarieties}*{3.17}.
 Since $\Fq$ is a finite field and hence perfect, $G^0$ is actually reduced over $\Fq$
 \cite{EGAIV2}*{Prop 6.4.1}.  Since every group scheme over a field is separated
 \cite{vdGeer-Moonen:AbelianVarieties}*{3.12},
 it follows that $G^0$ is a connected algebraic group.
\end{proof}

@@

\begin{lemma}\label{lem:extension}
The group homomorphism $\pi_0^*: \QCiso{\pi_0(G)} \to \QCiso{G}$ is injective.
\end{lemma}
\begin{proof}
Let $\qcs{L}$ be a quasicharacter sheaf on $\pi_0(G)$ and let $\rho : \pi_0^*\qcs{L} \to (\EE)_{G}$ be an isomorphism in $\QC(G)$. 
For each $x\in \pi_0(\bG)$, set $\bG^x \ceq \pi_0^{-1}(x)$.
The restriction $\pi_0^*\gqcs{L}\vert_{\bG^x}$ is the constant sheaf $(\gqcs{L}_x)_{\bG^x}$ so the isomorphism $\brho\vert_{\bG^x} : (\gqcs{L}_x)_{\bG^x} \to (\EE)_{\bG^x}$ determines an isomorphism $\brho_x : \gqcs{L}_x \to (\EE)_x$. The collection $\{ \brho_x \tq x\in \pi_0(\bG) \}$ determines an isomorphism $\qcs{L} \to (\EE)_{\pi_0(G)}$  in $\QC(\pi_0(G))$.
\end{proof}

@@

\begin{remark}
Although $\TrFrob{\pi_0(G)}$ is split and $\TrFrob{G^0}$ is an isomorphism, it does not follow that $\TrFrob{G}$ is split. 
\end{remark}

@@

In the opposite direction, let $\Nm : G(k') \to G(k)$ be the norm map and consider the group homomorphism:
\begin{align*}
\Nm^* : G(k)^* &\to G(k')^* \\
\chi &\mapsto \chi\circ \Nm.
\end{align*}
We can also interpret this operation in terms of quasicharacter sheaves.

If $\qcs{L} \ceq (\gqcs{L}, \mu, \phi)$ is a quasicharacter sheaf on $G$, we define
%a quasicharacter sheaf 
$\qcs{L}' \ceq (\gqcs{L}, \mu, \phi_{k'})$ on the base change
$G_{k'}$ of $G$ to $k'$ by setting
\[
\phi_{k'} \ceq \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi).
\]
The commutativity of the diagram (QC.3) for $\phi_{k'}$
follows from the fact that $\Frob{G_{k'}} = \Frob{G}^n$.
Note that we may also think about the construction of $\phi_{k'}$ from $\phi$
as restricting the action $\varphi$ of $\Weil{k}$ on $\gqcs{L}$,
defined in Section~\ref{ssec:category}, to $\Weil{k'}$.
%, $\Weil{k'}$.

@@

\begin{proof}
Let $\qcs{L} \ceq (\gqcs{L}, \mu, \phi) \in \QC(G)$ and write $F$ for $\Frob{G}$.  For any $x \in G(k')$,
we may compute the value of $\TrFrob{G_{k'}}(\nu_{k'/k}\qcs{L})(x)= t_{\nu_{k'/k}\qcs{L}}(x)$ as the trace of $\phi_{k'}$ on $\gqcs{L}_x$,
and the value of $\Nm^*(\TrFrob{G}(\qcs{L}))(x)$ as the trace of $\phi$ on $\gqcs{L}_{\Nm(x)}$.
Applying \ref{QC.3} to the stalk of $\gqcs{L}^{\boxtimes n}$ at the point $(x, \Frob{}(x), \ldots, \Frob{}^{n-1}(x))$ yields a diagram
\[
\begin{tikzcd}
\gqcs{L}_{\Nm(x)} \rar \dar{\phi_{\Nm(x)}} & \gqcs{L}_{F(x)} \otimes \gqcs{L}_{F^2(x)} \otimes \cdots \otimes \gqcs{L}_x
\dar{\phi_x \otimes (F^*\phi)_x \otimes \cdots \otimes ((F^{n-1})^*\phi)_x} \\
\gqcs{L}_{\Nm(x)} \rar & \gqcs{L}_x \otimes \gqcs{L}_{F(x)} \otimes \cdots \otimes \gqcs{L}_{F^{n-1}(x)}.
\end{tikzcd}
\]
Choose a basis vector $v$ for $\gqcs{L}_{\Nm(x)}$ and write $v_0 \otimes v_1 \otimes \cdots \otimes v_{n-1}$ for the image of $v$ under the
bottom map,
for $v_i \in \gqcs{L}_{\Frob{}^i(x)}$.  By \ref{QC.2}, $v$ maps to
$v_1 \otimes v_2 \otimes \cdots \otimes v_0$ along the top of the diagram.
Let $\alpha_i \in \EEx$ represent $((F^i)^*\phi)_x$ with respect to these bases and let $\alpha$ be
the trace of $\phi_{\Nm(x)}$.  We may now equate the trace $\alpha$ of $\phi$ on $\gqcs{L}_{\Nm(x)}$
with the product $\alpha_0 \cdots \alpha_{n-1}$, which is the trace of $\phi_{k'}$ on $\gqcs{L}_x$.
\end{proof}

Finally, let $G'$ be a smooth commutative group scheme over $k'$.
We explain how to geometrize the canonical isomorphism between characters of $G'(k')$ and of $(\Res_{k'/k}G')(k)$.
We may decompose the base change $(\Res_{k'/k}G')_{k'}$ of $\Res_{k'/k}G'$ to $k'$
into a product of copies of $G'$, indexed by elements of $\Gal(k'/k)$:
\[
(\Res_{k'/k}G')_{k'} \cong \prod_{\Gal(k'/k)} G'.
\]
Since products and coproducts agree for group schemes we have a natural inclusion of $k'$-schemes
\[
G' \hookrightarrow (\Res_{k'/k}G')_{k'},
\]
mapping $G'$ into the summand corresponding to $1 \in \Gal(k'/k)$.  Composing $\nu_{k'/k}$
from Proposition~\ref{prop:basechange} with pullback along this map yields a functor
\[
\rho : \QC(\Res_{k'/k}G') \to \QC(G').
\]

\begin{proposition}
Let $k'/k$ be a finite extension and let $G'$ be a smooth commutative group scheme over $k'$.
Then the functor 
\[
\rho : \QC(\Res_{k'/k}G') \to \QC(G'),
\]
defined above, induces
\[
\begin{tikzcd}
\QCiso{\Res_{k'/k} G'} \dar{\TrFrob{\Res_{k'/k} G'}} \rar{\rho} & \QCiso{G'} \dar{\TrFrob{G'}}\\
G'(k')^* \rar & G'(k')^*,
\end{tikzcd}
\]
where the bottom map is the identity.
\end{proposition}
\begin{proof}
By Lemma~\ref{lem:pullback} the pullback part of the definition of $\rho$ corresponds to the map
\[
(\Res_{k'/k}G')(k')^* \to G'(k')^*
\]
induced by $g \mapsto (g, 1, \ldots, 1)$.  Since the action of $\Gal(k'/k)$ on
\[
(\Res_{k'/k}G')_{k'} \cong \prod_{\Gal(k'/k)} G'
\]
is given by permuting coordinates, composition with the norm map yields the identity on $G'(k')$.
\end{proof}
